Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-12-11 manuscripts/manuscript.Rmd
pdf figures

```{r figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.  Curves show the expectations over the posterior step-ahead predictions based on each of the estimated models. Observe that all models are fitting the data reasonably well.)", fig.width=8, fig.height=6, cache=FALSE}
step_ahead_posteriors <- function(x){
  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
  df_post <- melt(lapply(sample(100, 30),  
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}
df_post <- step_ahead_posteriors(x)
library("dplyr")
df_post_means <- summarize(group_by(df_post, time,variable,stock), value = mean(value))
ggplot(df_post_means,aes(time, value)) +
  geom_point(aes(time, stock), size=2) + 
  geom_line(aes(time, value, col=variable, group=variable),lwd=1) + 
#  geom_line(aes(time, value, col=variable, group=interaction(L1,variable))) + 
  facet_wrap(~variable) + 
  scale_colour_manual(values=colorkey) +  
  theme(legend.position="none")

@@

```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))
pandoc.table(rbind(dictable, aictable, bictable))
```

@@

```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated (MLE) process.  We show the expected value for the function $f$ under each model.  Plots show replicates drawn from the posterior distributions in order to convey uncertainty of the estimates.  Note the MLE is a point estimate of parameters and so reflects no uncertainty in the distribution. The training data are also shown as black points.  The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6, cache=FALSE, dev="CairoJPEG", dev.args=c(quality=100)}
# subset data first, don't care about seeing dynamics for stock >> K
x_grid_short <- x_grid[1:max(which(x_grid<12))]
gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)

@@

```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discretization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6, cache=FALSE}
policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
names(policies) <- c("stock", "method", "value")

@@

```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models. We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE, dev="CairoJPEG", dev.args=c(quality=100)}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
  scale_colour_manual(values=colorkey, guide = FALSE) +
  xlab("stock size, x(t)") +
  facet_wrap(~method) + 
  guides(legend.position="none", color=FALSE)

>>>>

```{r figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.  Curves show the expectations over the posterior step-ahead predictions based on each of the estimated models. Observe that all models are fitting the data reasonably well.)", fig.width=8, fig.height=6}
step_ahead_posteriors <- function(x){
  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
  df_post <- melt(lapply(sample(100, 30),  
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}
df_post <- step_ahead_posteriors(x)
library("dplyr")
df_post_means <- summarize(group_by(df_post, time,variable,stock), value = mean(value))
ggplot(df_post_means,aes(time, value)) +
  geom_point(aes(time, stock), size=2) + 
  geom_line(aes(time, value, col=variable, group=variable),lwd=1) + 
#  geom_line(aes(time, value, col=variable, group=interaction(L1,variable))) + 
  facet_wrap(~variable) + 
  scale_colour_manual(values=colorkey) +  
  theme(legend.position="none")

@@

```{r deviances, dependson=c("dic_calc"), include = FALSE, echo = FALSE}
allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))
pandoc.table(rbind(dictable, aictable, bictable))
```

@@

```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated (MLE) process.  We show the expected value for the function $f$ under each model.  Plots show replicates drawn from the posterior distributions in order to convey uncertainty of the estimates.  Note the MLE is a point estimate of parameters and so reflects no uncertainty in the distribution. The training data are also shown as black points.  The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
# subset data first, don't care about seeing dynamics for stock >> K
x_grid_short <- x_grid[1:max(which(x_grid<12))]
gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)

@@

```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discretization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6}
policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
names(policies) <- c("stock", "method", "value")

@@

```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models. We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
  scale_colour_manual(values=colorkey, guide = FALSE) +
  xlab("stock size, x(t)") +
  facet_wrap(~method) + 
  guides(legend.position="none", color=FALSE)
